---
title       : Age Calculator Project
subtitle    : Shiny project for Developing Data Products course
author      : Jalal Eddin Omari
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
knit        : slidify::knit2slides
---

## Project Idea


Project idea is to host a shiny application to calculate user age.

User has to pick his birth date from date picker and this program will display his age.

Also application will show a simple age plot based on year.

You can test the application by clicking [here](http://jalalomary.shinyapps.io/shiny).

--- .smallCode

## Calculation Code
<style>
.smallCode pre{
  font-size:12px;
}
</style>
```{r}
age_years <- function(earlier, later)
{
  lt <- data.frame(earlier, later)
  age <- as.numeric(format(lt[,2],format="%Y")) - as.numeric(format(lt[,1],format="%Y"))
  
  dayOnLaterYear <- ifelse(format(lt[,1],format="%m-%d")!="02-29",
                           as.Date(paste(format(lt[,2],format="%Y"),"-",format(lt[,1],format="%m-%d"),sep="")),
                           ifelse(as.numeric(format(later,format="%Y")) %% 400 == 0 | as.numeric(format(later,format="%Y")) %% 100 != 0 &
                           as.numeric(format(later,format="%Y")) %% 4 == 0,
                                  as.Date(paste(format(lt[,2],format="%Y"),"-",format(lt[,1],format="%m-%d"),sep="")),
                                  as.Date(paste(format(lt[,2],format="%Y"),"-","02-28",sep=""))))
  
  age[which(dayOnLaterYear > lt$later)] <- age[which(dayOnLaterYear > lt$later)] - 1
  
  age
}
```
Calling Live Example

```{r}
age_years(as.Date("1990-02-26"),Sys.Date())
```

--- .smallCode

## ui.R
```{r eval=FALSE}
library(shiny)

shinyUI(pageWithSidebar(

  headerPanel("Age Calculator"),

  sidebarPanel(
    helpText("This simple application will calculate your age, 
             you can fill your birth date below, 
             and it will automatically populate the calculated 
             age and display it on the right side."),
    dateInput("dob", "Birth Date:", max = Sys.Date() -1, value="1990-01-01")
  ),

  mainPanel(
    h3("Your age is: "),
    verbatimTextOutput("age"),
    plotOutput("plot")
  )
))
```

--- .smallCode

## server.R
```{r eval=FALSE}
library(shiny)

source("age_function.R")

shinyServer(function(input, output) {
    age <- reactive({age_years(input$dob,Sys.Date())})
    output$age <- renderText({age()})
    start_year <- reactive({as.numeric(format(input$dob, "%Y"))})
    end_year <- reactive({start_year() + age()})
    output$plot <- renderPlot({
      plot(c(start_year():end_year()), c(0:age()), xlab="Year", ylab="Age",
           lines(c(start_year():end_year()), c(0:age()),col="red"))
      
    })
})
```

### References:

Age Calculation Function by: nzcoops, [http://www.r-bloggers.com/updated-age-calculation-function/](http://www.r-bloggers.com/updated-age-calculation-function/)
